---
title: "Lesson 9: Diamonds & Price Predictions"
author: "Dannyel Cardoso da Fonseca"  
date: "`r format(Sys.Date())`"
output: github_document
---

```{r echo=FALSE}
setwd("~/Documents/GitHub/udacity-dand/core-course-05-exploratory-data-analysis/lesson-09-diamonds-and-price-predictions")
```

### Load Libraries and Datasets

```{r Load Libraries and Datasets}
library(ggplot2)
library(gridExtra, warn.conflicts = FALSE)
theme_set(theme_light())

data(diamonds)
```

### Welcome!
> Note: https://www.youtube.com/watch?v=KJT4Z0xpHns

### Linear Regression Models
> Note: In this lesson, Solomon will be using a linear regression model to predict diamond price using other variables in the diamonds dataset. If you are not familiar with linear regression, you may want to take a break and go through [Lesson 18 of Udacity's Intro to Inferential Statistics](https://classroom.udacity.com/courses/ud201/lessons/1309228537/concepts/1822138740923) course, which covers linear regression. When you're done, you'll be ready to come back and apply your new knowledge to the diamonds dataset!

### Scatterplot Review
> Note: https://www.youtube.com/watch?v=W96zaGEma7o

**Quiz:** Let's start by examining two variables in the data set. The scatterplot is a powerful tool to help you understand the relationship between two continuous variables.

We can quickly see if the relationship is linear or not. In this case, we can use a variety of diamond characteristics to help us figure out whether the price advertised for any given diamond is reasonable or a rip-off.

Let's consider the price of a diamond and it's carat weight. Create a scatterplot of price (y) vs carat weight (x).

Limit the x-axis and y-axis to omit the top 1% of values.

**Response:** 
```{r Scatterplot Review}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 1/10, na.rm = TRUE) +
  scale_x_continuous(limits = c(0, quantile(diamonds$carat, .99)),
                     breaks = seq(0, 2.5, 0.25)) +
  scale_y_continuous(limits = c(0, quantile(diamonds$price, .99)),
                     breaks = seq(0, 20000, 1500))
```

### Price and Carat Relationship
> Note: https://www.youtube.com/watch?v=gG4xwgj1yVA

**Quiz:** What do you notice about the relationship between price and carat?

**Response:** Apparently, as carat increases the price increase as well, in a non-linear way. As carat increases price dispersion increases as well.
```{r Price and Carat Relationship}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point(alpha = 1/10, na.rm = TRUE) +
  geom_smooth(method = "lm", na.rm = TRUE) +
  scale_x_continuous(limits = c(0, quantile(diamonds$carat, .99)),
                     breaks = seq(0, 2.5, 0.25)) +
  scale_y_continuous(limits = c(0, quantile(diamonds$price, .99)),
                     breaks = seq(0, 20000, 1500))
```

### Frances Gerety
> Note: https://www.youtube.com/watch?v=GXT_vXBA0vQ

**Quiz:** Frances Gerety coined a famous slogan.

**Response:** "A diamond is forever."




