---
title: "Lesson 8: Problem Set - Explore Many Variables"
author: "Dannyel Cardoso da Fonseca"  
date: "`r format(Sys.Date())`"
output: github_document
---

```{r echo=FALSE}
setwd("~/Documents/GitHub/udacity-dand/core-course-05-exploratory-data-analysis/lesson-08-problem-set-explore-many-variables")
```

### Load Libraries and Datasets

```{r Load Libraries and Datasets}
library(ggplot2)

data(diamonds)
```

### Price Histograms with Facet and Color

**Quiz:** Create a histogram of diamond prices. Facet the histogram by diamond color and use cut to color the histogram bars.

The plot should look something like this: http://i.imgur.com/b5xyrOu.jpg

> Note: In the link, a color palette of type 'qual' was used to color the histogram using scale_fill_brewer(type = 'qual')

**Response:**
```{r Price Histograms with Facet and Color}
ggplot(diamonds, aes(x = price, fill = cut)) +
  geom_histogram(bins = 30) +
  facet_wrap(~color) +
  scale_x_log10() +
  scale_fill_brewer(type = 'qual')
```

### Price vs. Table Colored by Cut

**Quiz:** Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond.

The plot should look something like this: http://i.imgur.com/rQF9jQr.jpg

> Note: In the link, a color palette of type 'qual' was used to color the scatterplot using scale_color_brewer(type = 'qual')

**Response:**
```{r Price vs. Table Colored by Cut}
ggplot(diamonds, aes(x = table, y = price)) +
  geom_point(aes(color = cut), position = position_jitter(width = .1), alpha = .3, na.rm = TRUE) +
  scale_x_continuous(limits = c(50, 80), breaks = seq(50, 80, 2)) +
  scale_y_continuous(breaks = seq(0, 20000, 1000)) +
  scale_fill_brewer(type = 'qual')
```

### Typical Table Value

**Quiz:** What is the typical table range for the majority of diamonds of ideal cut?

**Response:** 53 to 57.

**Quiz:** What is the typical table range for the majority of diamonds of premium cut?

**Response:** 58 to 61.


