---
title: "Lesson 7: Explore Many Variables"
author: "Dannyel Cardoso da Fonseca"  
date: "`r format(Sys.Date())`"
output: github_document
---

```{r echo=FALSE}
setwd("~/Documents/GitHub/udacity-dand/core-course-05-exploratory-data-analysis/lesson-07-explore-many-variables")
```

### Load Libraries and Datasets

```{r Load Libraries and Datasets}
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)

pf <- read.delim('lesson_07_files/data/pseudo_facebook.tsv')
yo <- read.csv("lesson_07_files/data/yogurt.csv")
yo$id <- as.factor(yo$id)
```
### Multivariate Data
> Note: https://www.youtube.com/watch?v=jsg6lhrJN1g

### Perceived Audience Size By Age
> Note: https://www.youtube.com/watch?v=GFKRNBnFGVU

### Third Qualitative Variable
> Note: https://www.youtube.com/watch?v=Q2M8xyY47fc

**Quiz:** Write code to create a new data frame, called 'pf.fc_by_age_gender', that contains information on each age AND gender group.

The data frame should contain the following variables:

    mean_friend_count,
    median_friend_count,
    n (the number of users in each age and gender grouping)

**Response:**
```{r Third Qualitative Variable}
pf.fc_by_age_gender <- pf %>% 
                          filter(!is.na(gender)) %>% 
                          group_by(age, gender) %>% 
                          summarise(mean_friend_count = mean(friend_count),
                                    median_friend_count = as.integer(median(friend_count)),
                                    n = n()) %>%                   
                          ungroup() %>% 
                          arrange(age, gender)

head(pf.fc_by_age_gender)
```

### Plotting Conditional Summaries
> Note: https://www.youtube.com/watch?v=8SqL0v_FSsc

**Quiz:** Create a line graph showing the median friend count over the ages for each gender. Be sure to use the data frame you just created, pf.fc_by_age_gender.

**Response:**
```{r Plotting Conditional Summaries}
ggplot(pf.fc_by_age_gender, aes(x = age, y = mean_friend_count)) +
  geom_line(aes(color = gender)) +
  scale_x_continuous(breaks = seq(10, 115, 5)) +
  scale_y_continuous(breaks = seq(0, 750, 50))
```

### Thinking In Ratios
> Note: https://www.youtube.com/watch?v=r4ZOwz3_oXs

### Wide And Long Format
> Note: https://www.youtube.com/watch?v=zlaeISxRESQ

### Reshaping Data
> Note: https://www.youtube.com/watch?v=zQj_waidR5w

```{r Reshaping Data}
library(tidyr)
pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>% 
                               select(age, gender, median_friend_count) %>% 
                               spread(gender, median_friend_count)

head(pf.fc_by_age_gender.wide)
```

### Ratio Plot
> Note: https://www.youtube.com/watch?v=gfZ7C-QBF0k

**Quiz:** Plot the ratio of the female to male median friend counts using the data frame pf.fc_by_age_gender.wide.

Think about what geom you should use. Add a horizontal line to the plot with a y intercept of 1, which will be the base line. Look up the documentation for geom_hline to do that. Use the parameter linetype in geom_hline to make the line dashed.

**Response:**
```{r Ratio Plot}
ggplot(pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) +
  geom_line() +
  geom_hline(yintercept = 1, linetype = 2, alpha = .3) +
  scale_x_continuous(breaks = seq(10, 115, 5)) +
  scale_y_continuous(breaks = seq(0, 4.5, .25))
```

### Third Quantitative Variable
> Note: https://www.youtube.com/watch?v=gpwlI9Wa8xI

**Quiz:** Create a variable called year_joined in the pf data frame using the variable tenure and 2014 as the reference year.

The variable year joined should contain the year that a user joined facebook.

**Response:**
```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling(pf$tenure/365)

pf.fc_by_year_joined <- pf %>% 
                          filter(!is.na(year_joined)) %>% 
                          group_by(year_joined) %>% 
                          summarise(count = n())

ggplot(pf.fc_by_year_joined, aes(x = year_joined, y = count)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -1, size=3) +
  scale_x_continuous(breaks = seq(2005, 2014, 1)) +
  scale_y_continuous(breaks = seq(0, 50000, 3000))
```

### Cut a Variable
> Note: https://www.youtube.com/watch?v=n0lluEhKUfQ

**Quiz:** Create a new variable in the data frame called year_joined.bucket by using the cut function on the variable year_joined.

You need to create the following buckets for the new variable, year_joined.bucket

    (2004, 2009]
    (2009, 2011]
    (2011, 2012]
    (2012, 2014]

Note that a parenthesis means exclude the year and a bracket means include the year.

**Response:**
```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))

unique(pf$year_joined.bucket)
```

### Plotting It All Together
> Note: https://www.youtube.com/watch?v=CiS4rBbr6tw

**Quiz:** Create a line graph of friend_count vs. age so that each year_joined.bucket is a line tracking the median user friend_count across age. This means you should have four different lines on your plot.

You should subset the data to exclude the users whose year_joined.bucket is NA.

**Response:**
```{r Plotting It All Together}
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = median) +
  scale_x_continuous(breaks = seq(10, 120, 10)) +
  scale_y_continuous(breaks = seq(0, 2000, 100))
```

### Plot the Grand Mean
> Note: https://www.youtube.com/watch?v=pxaXkCjukGM

**Quiz:** Write code to do the following:

(1) Add another geom_line to code below to plot the grand mean of the friend count vs age.

(2) Exclude any users whose year_joined.bucket is NA.

(3) Use a different line type for the grand mean.

> As a reminder, the parameter linetype can take the values 0-6: 0 = blank, 1 = solid, 2 = dashed 3 = dotted, 4 = dotdash, 5 = longdash 6 = twodash

**Response:**
```{r Plot the Grand Mean}
ggplot(subset(pf, !is.na(year_joined.bucket)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = mean, linetype = 5) + # grand mean
  scale_x_continuous(breaks = seq(10, 120, 10)) +
  scale_y_continuous(breaks = seq(0, 2000, 100))
```

### Friending Rate
> Note: https://www.youtube.com/watch?v=ZO7y9tsSQ0A

**Quiz:** Calculate how many friends does a user have for each day since they have started using the service.

What is the median friend rate?

What is the maximum friend rate?

**Response:**
```{r Friending Rate}
with(subset(pf, tenure > 0), summary(friend_count/tenure))

ggplot(subset(pf, tenure > 0), aes(y = friend_count/tenure, x = "")) +
  geom_boxplot() +
  geom_point(stat = "summary", fun.y = mean, color = "red") +
  coord_cartesian(ylim = c(0, 1.5)) +
  scale_y_continuous(breaks = seq(0, 1.5, .1)) +
  xlab("")
```

### Friendships Initiated
> Note: https://www.youtube.com/watch?v=Keh5GwaSWdk

**Quiz:** Create a line graph of mean of friendships_initiated per day (of tenure) vs. tenure colored by year_joined.bucket.

You need to make use of the variables tenure, friendships_initiated, and year_joined.bucket.

You also need to subset the data to only consider user with at least one day of tenure.

**Response:**
```{r Friendships Initiated}
ggplot(subset(pf, tenure > 0), aes(x = tenure, y = friendships_initiated/tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
  scale_x_continuous(breaks = seq(0, 3500, 250)) +
  scale_y_continuous(breaks = seq(0, 10, .5))
```

### Bias Variance Trade off Revisited
> Note: https://www.youtube.com/watch?v=CSVf96g0XGM

**Quiz:** Instead of geom_line(), use geom_smooth() to add a smoother to the plot. You can use the defaults for geom_smooth() but do color the line by year_joined.bucket

**Response:**
```{r Bias Variance Trade off Revisited}
library(gridExtra, warn.conflicts = FALSE)

p1 <- ggplot(subset(pf, tenure > 0), aes(x = 7 * round(tenure / 7), y = friendships_initiated/tenure)) +
        geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
        scale_x_continuous(breaks = seq(0, 3500, 250)) +
        scale_y_continuous(breaks = seq(0, 10, 1.5)) +
        ylab("fs_initiated/tenure")

p2 <- ggplot(subset(pf, tenure > 0), aes(x = 30 * round(tenure / 30), y = friendships_initiated/tenure)) +
        geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
        scale_x_continuous(breaks = seq(0, 3500, 250)) +
        scale_y_continuous(breaks = seq(0, 10, 1)) +
        ylab("fs_initiated/tenure")

p3 <- ggplot(subset(pf, tenure > 0), aes(x = 90 * round(tenure / 90), y = friendships_initiated/tenure)) +
        geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean) +
        scale_x_continuous(breaks = seq(0, 3500, 250)) +
        scale_y_continuous(breaks = seq(0, 10, .5)) +
        ylab("fs_initiated/tenure")

grid.arrange(p1, p2, p3, ncol = 1)

ggplot(subset(pf, tenure > 0), aes(x = 7 * round(tenure / 7), y = friendships_initiated /tenure)) +
  geom_smooth(aes(color = year_joined.bucket)) +
  scale_x_continuous(breaks = seq(0, 3500, 250)) +
  scale_y_continuous(breaks = seq(0, 4, .5)) 
```

### Sean's NFL Fan Sentiment Study
> Note: https://www.youtube.com/watch?v=ahaxt6UKxQw

### Introducing The Yogurt Dataset
> Note: https://www.youtube.com/watch?v=5J9GxnJVo78

### Histograms Revisited
> Note: https://www.youtube.com/watch?v=7PyV7HxpSYA

**Quiz:** Create a histogram of yogurt prices.

Use the qplot or ggplot syntax. Don't add any extra code for labels, titles, or colors. We're just looking for the basic syntax here.

**Response:**
```{r Histograms Revisited 1}
ggplot(yo, aes(x = price)) +
  geom_histogram()
```

I can note that the amount of yogurt increases as the price of yogurt increases. But the growth of yogurt quantity is not linear. It is intercalated between very low quantity and high quantity as the price increases.

```{r Histograms Revisited 2}
ggplot(yo, aes(x = price)) +
  geom_histogram(binwidth = 10)
```

Now, I can note that the amount of yogurt increases as the price of yogurt increases in an exponential way. 


