---
output: github_document
---
EDA in Issues Tracking Data Set
================================================================================
*by Dannyel Cardoso da Fonseca* 

```{r echo=FALSE, message=FALSE, warning=FALSE, "Load Packages"}
library(ggplot2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, "Set Up"}
# NOTE: To run this notebook set up current work directory to this file location

# Set up default theme for plots
theme_set(theme_classic())
```

```{r echo=FALSE, "Load Data Set"}
# Load the Issues Tracking Data Set
issues <- read.csv("issues_tracking.csv")
```

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots}
ggplot(issues, aes(x = issue_type, y = ..count..)) +
    geom_bar()
```